Modelling state-switching functional data with hidden Markov models: reviewer 2 response

### 2021-07-09 Report/Reviewer_2.tex
Update page / line numbers, reviewer 2

    We thank the reviewer for the references and have added them to page 5, lines 92-96.

@@

    It is confusing that the number of curves is denoted by $T$, which usually indicates time points. However, the curves are assumed to be sequential in time, and there is a connection between the coarse scale index $t$ and the fine-scale index $t^*$. We have added a sentence on page 7, lines 137-138 to clarify that these curves are sequential but not necessarily equi-spaced in time. In addition, there is some precedence for indexing sequential curves with $t$. See ``Cam\'er-Karhunen-Lo\`eve representation and harmonic principal component analysis of functional time series" (2013) by Panaretos and Tavakoli and ``Dynamic functional principal components" by Hormann and Kidinski (2014) for examples.

@@

    We agree and have added a sentence to clarify this on page 8, lines 151-152. Thank you for noting the omission.

@@

    We have edited page 9, lines 171-176 to clarify the difference between the likelihood and the forward algorithm. We have also described the forward algorithm, which is used to evaluate the likelihood, as well as the forward-backward algorithm, which is used for state decoding and maximizing the likelihood via the Baum-Welch algorithm (Zucchini et al, 2016).

@@

    We personally maximized the likelihood directly. For each model, we ran 100 direct maximization routines using random initializations using the Nelder-Mead method of the optimize package of SciPy. In particular, we optimized one row at a time of the probability transition matrices and one set of parameters associated with a (feature, hidden state) pair at a time. We found that most, but not all, of the optimization routines converged to similar parameter estimates. We have added an explanation of our optimization routine on page 26, lines 491-500. Thank you for calling attention to this omission.

@@

    This is a very useful and common practice in the ecology literature. While it is beyond the scope of our case study, we have added a brief discussion of covariates on page 10, lines 182-185.

>>>>

    We thank the reviewer for the references and have added them to page 5, lines 92-95.

@@

    It is confusing that the number of curves is denoted by $T$, which usually indicates time points. However, the curves are assumed to be sequential in time, and there is a connection between the coarse scale index $t$ and the fine-scale index $t^*$. We have added a sentence on page 7, lines 138-140 to clarify that these curves are sequential but not necessarily equi-spaced in time. In addition, there is some precedence for indexing sequential curves with $t$. See ``Cam\'er-Karhunen-Lo\`eve representation and harmonic principal component analysis of functional time series" (2013) by Panaretos and Tavakoli and ``Dynamic functional principal components" by Hormann and Kidinski (2014) for examples.

@@

    We agree and have added a sentence to clarify this on page 8, lines 153-154. Thank you for noting the omission.

@@

    We have edited pages 9, lines 165-172 to clarify the difference between the likelihood and the forward algorithm. We have also described the forward algorithm, which is used to evaluate the likelihood, as well as the forward-backward algorithm, which is used for state decoding and maximizing the likelihood via the Baum-Welch algorithm (Zucchini et al, 2016).

@@

    We personally maximized the likelihood directly. For each model, we ran 100 direct maximization routines using random initializations using the Nelder-Mead method of the optimize package of SciPy. In particular, we optimized one row at a time of the probability transition matrices and one set of parameters associated with a (feature, hidden state) pair at a time. We found that most, but not all, of the optimization routines converged to similar parameter estimates. We have added an explanation of our optimization routine on page 26, lines 491-500. Thank you for prompting us to correct this omission.

@@

    This is a very useful and common practice in the ecology literature. While it is beyond the scope of our case study, we have added a brief discussion of covariates on page 10, lines 183-186.
